### 2020-01-21 thesis/pretext.tex

%\thispagestyle{plain}
\thispagestyle{empty}
\begin{center}
	\large
	AN ABSTRACT OF THE THESIS OF
\end{center}

\justify{}
\underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
presented on \underline{\jpsays{include the correct date here!}}.
\vspace{1.0cm}

\justify{}
Title: \underline{\jpsays{Title of the thesis}}.
\vspace{2.0cm}

\justify{}
Abstract approved:
\underline{\hspace{11cm}}
\vspace{0.5cm}

%\centering{}
\hspace{7.0cm} David J. Roundy
\vspace{1.0cm}

\justify{}
\doublespacing
"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip
ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore 
eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia
deserunt mollit anim id est laborum."

\iffalse
\singlespacing
"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt..."
\fi

\newpage{}
\thispagestyle{empty}
\singlespacing

\vspace*{4.0cm}
\begin{center}
$\copyright$ Copyright by Jordan K. Pommerenck \\
\jpsays{include the correct date here!} \\
(All Rights Reserved)
\end{center}

\newpage{}
\thispagestyle{empty}
\singlespacing

\begin{center}
\jpsays{Title of the thesis}

\vspace{1.0cm}
by \\
Jordan K. Pommerenck \\
\vspace{3.0cm}
A DISSERTATION \\
\vspace{0.5cm}
submitted to \\
\vspace{0.5cm}
Oregon State University \\
\vspace{3.0cm}
in partial fulfillment of \\
the requirements for the \\
degree of \\
\vspace{1.0cm}
Doctor of Philosophy \\
\vspace{3.0cm}
Presented October 26, 2019  \\
Commencement December 2019 \jpsays{include the correct date here!}
\end{center}

\newpage{}
\thispagestyle{empty}
\singlespacing
\justify{}
\underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
\jpsays{get the date right here? August 26, 2015}. \\

\justify{}
\vspace{0.5cm}
APPROVED: \\

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Major Professor, representing Physics
\vspace{1.0cm}
\end{flushleft}

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Chair of the Department of Physics
\vspace{1.0cm}
\end{flushleft}

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Dean of the Graduate School
\vspace{1.0cm}
\end{flushleft}

\vspace{3.0cm}
\justify{}
I understand that my dissertation will become part of the permanent collection
of Oregon State University libraries. My signature below authorizes release of my
dissertation to any reader upon request. \\
\vspace{0.5cm}
\begin{flushleft}
\underline{\hspace{15cm}}
\end{flushleft}
\centering{}
Jordan K. Pommerenck, Author

\newpage{}
\thispagestyle{empty}
\begin{center}
	\large
	ACKNOWLEDGEMENTS
\end{center}
\justify{}
\doublespacing
"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
dolore magna aliqua. Ut