### 2020-02-17 thesis/pretext.tex
added abstract and some data.

%\thispagestyle{plain}
\thispagestyle{empty}
\begin{center}
	\large
	AN ABSTRACT OF THE THESIS OF
\end{center}

\justify{}
\underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
presented on \underline{\jpsays{include the correct date here FORMAT = April 10, 2020!}}.
\vspace{1.0cm}

\justify{}
Title: \underline{\jpsays{Title of the thesis}}.
\vspace{2.0cm}

\justify{}
Abstract approved:
\underline{\hspace{11cm}}
\vspace{0.5cm}

%\centering{}
\hspace{7.0cm} David J. Roundy
\vspace{1.0cm}

\justify{}
\doublespacing
This work introduces the development of the novel flat-histogram Monte 
Carlo (MC) method SAD and explores the convergence properties of a 
variety of related weight-based MC methods. The new method is applied 
to a variety of physical `test' systems including the 2D Ising model, 
square-well fluid, and Lennard-Jones clusters. A driving motivation for 
developing novel Monte Carlo methods that have physically based tunable 
parameters is to provide simulation methods for exploring metal-organic 
frameworks (MOFs). A theoretical framework for gas adsorption and 
delivery in metal-organic frameworks is developed and compared with 
experimental data. SAD is a introduced as a powerful method for 
examining thermodynamic properties for MOFs.  A \emph{preliminary} 
groundwork is laid for the future development of 2D SAD which would be 
able to compute isotherms for any given MOF thereby providing a 
powerful way for researchers to determine a MOF's suitability for gas 
storage applications.

\newpage{}
\thispagestyle{empty}
\singlespacing

\vspace*{4.0cm}
\begin{center}
$\copyright$ Copyright by Jordan K. Pommerenck \\
\jpsays{include the correct date here!} \\
(All Rights Reserved)
\end{center}

\newpage{}
\thispagestyle{empty}
\singlespacing

\begin{center}
\jpsays{Title of the thesis}

\vspace{1.0cm}
by \\
Jordan K. Pommerenck \\
\vspace{3.0cm}
A DISSERTATION \\
\vspace{0.5cm}
submitted to \\
\vspace{0.5cm}
Oregon State University \\
\vspace{3.0cm}
in partial fulfillment of \\
the requirements for the \\
degree of \\
\vspace{1.0cm}
Doctor of Philosophy \\
\vspace{3.0cm}
Presented April 10, 2020  \\
Commencement June 2020 \jpsays{include the correct date here!}
\end{center}

\newpage{}
\thispagestyle{empty}
\singlespacing
\justify{}
\underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
\jpsays{get the date right here? April 10, 2015}. \\

\justify{}
\vspace{0.5cm}
APPROVED: \\

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Major Professor, representing Physics
\vspace{1.0cm}
\end{flushleft}

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Chair of the Department of Physics
\vspace{1.0cm}
\end{flushleft}

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Dean of the Graduate School
\vspace{1.0cm}
\end{flushleft}

\vspace{3.0cm}
\justify{}
I understand that my dissertation will become part of the permanent collection
of Oregon State University libraries. My signature below authorizes release of my
dissertation to any reader upon request. \\
\vspace{0.5cm}
\begin{flushleft}
\underline{\hspace{15cm}}
\end{flushleft}
\centering{}
Jordan K. Pommerenck, Author

\newpage{}
\thispagestyle{empty}
\begin{center}
	\large
	ACKNOWLEDGEMENTS
\end{center}
\justify{}
\doublespacing
\jpsays{FIXME: "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
dolore magna aliqua. Ut}