### 2020-04-30 thesis/chapter-5.tex
added in final chapters including author contributions.

\chapter{General Conclusion}

Hello this is the conclusion.