### 2020-04-30 thesis/thesis.tex

\documentclass[12pt,a4paper]{report}
\usepackage{fancyhdr}

% for creating an index at end of document
\usepackage{imakeidx}
\makeindex[columns=1, title=Alphabetical Index]

% indexing commands
\newcommand{\bind}[1]{\textbf{#1}\index{#1}}
\newcommand{\iind}[1]{\textit{#1}\index{#1}}
\newcommand{\ind}[1]{#1\index{#1}}

% link to section command
\newcommand{\HR}[1]{\textbf{\hyperref[#1]{\ref{#1}}}}

\usepackage[hypertexnames=false]{hyperref}
\hypersetup{
    unicode=false,          % non-Latin characters in Acrobat’s bookmarks
    pdftoolbar=true,        % show Acrobat’s toolbar?
    pdfmenubar=true,        % show Acrobat’s menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={Pommerenck Thesis},    % title
    pdfauthor={Jordan K. Pommerenck},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={flat-histogram, Monte Carlo, Thesis}, % list of keywords
    pdfnewwindow=true,      % links in new PDF window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
    citecolor=black,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}

\pagestyle{fancy}
\fancyhf{}
\fancyheadoffset{0cm}
\renewcommand{\headrulewidth}{0pt} 
\renewcommand{\footrulewidth}{0pt}
\fancyhead[R]{\thepage}
\fancypagestyle{plain}{%
   \fancyhf{}%
   \fancyhead[R]{\thepage}%
}

\usepackage{blindtext}
\usepackage{parskip}

\usepackage[left=1.5in,right=1in,top=1in,bottom=1in,includefoot,heightrounded]{geometry}
%\usepackage[pagestyles]{titlesec}

%\sethead{\itshape\chaptername\thechapter}{}{\itshape\chaptertitle}
%\titleformat{\chapter}[display]{\normalfont\bfseries\Large\centering}{}{18pt}{\Large}
%\titlespacing{\chapter}{0pt}{0pt}{1cm}

%\usepackage{blindtext}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[document]{ragged2e}
% differential and partials package
\usepackage[thinc]{esdiff}
\usepackage{physics}
\newcommand{\D}[1]{\text{d}#1}
% the nth superscript package
\usepackage[super]{nth}

\usepackage{indentfirst}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{amsmath,amssymb}
\usepackage{nicefrac}

\usepackage[backend=bibtex,maxnames=200,sorting=none]{biblatex}
%\usepackage[square,numbers]{natbib}

%\addbibresource{thesis.bib}

% All the packages I have added...
\usepackage{setspace} % Allows double spacing with the \doublespacing command
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{\thepage}

\usepackage{color}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\red{Jordan:} \emph{#1}]}}

\author{Jordan K. Pommerenck}

% rename Contents to ToC and center
\usepackage{tocloft}
\renewcommand{\cftchapfont}{\normalfont}
\renewcommand{\cftchappagefont}{\normalfont}
\renewcommand{\contentsname}{\hfill\Large\normalfont TABLE OF CONTENTS}
\renewcommand{\cftaftertoctitle}{\hfill}
\renewcommand{\listfigurename}{\hfill\Large\normalfont LIST OF FIGURES\hfill}
\renewcommand{\cftaftertoctitle}{\hfill}

% THE BIBLIOGRAPHY NEEDS TO COME BEFORE ALL OF THE CHAPTERS
\bibliography{thesis}

\begin{document}
\input{pretext.tex}

\clearpage{}
\tableofcontents
\thispagestyle{empty}

\clearpage{}
\listoffigures
\thispagestyle{empty}

\clearpage{}

\setcounter{page}{1} % needed if pages are to be continuously numbered

% introduction
\input{chapter-1}
% sq-well paper
\input{chapter-2}
% ising paper
\input{chapter-3}
% gas-adsorption paper
\input{chapter-4}
% general conclusion linking all manuscripts
\input{chapter-5}

\printbibliography[title={Bibliography}]

\clearpage{}
%\printindex
\end{document}