### 2020-05-02 ising/ising.tex

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}
\usepackage{braket}

% define colors for comments
\definecolor{dark-gray}{gray}{0.10}
\definecolor{light-gray}{gray}{0.70}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
\begin{document}

\title{Flat-histogram method comparison on 2D Ising model
}

\author{Jordan K. Pommerenck} \author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
We examine the convergence of the stochastic approximation with a dynamic
update factor (SAD) algorithm applied to the 2D Ising model. Comparison with
stochastic approximation Monte Carlo (SAMC) and Wang-Landau (WL) methods show that SAD performs robustly and without user input knowledge of an energy range. We also compare SAD with ``production'' run WL with fixed weights which is the first of its kind recorded. Since SAD is more powerful in the common case in which the range of energies is not known in advance, the 2D Ising model presents a unique challenge since the energy range is readily known and easily discovered by WL methods.
\end{abstract}

\maketitle

\input{ising-body.tex}

\bibliography{ising} % Produces the bibliography via BibTeX.

\end{document}