### 2020-05-02 thesis/chapter-3.tex

\chapter{Flat-histogram method comparison on 2D Ising Model
}

\input{../ising/ising-body.tex}