### 2020-05-05 gas-adsorption/gas-ads.tex
update the paper and conform to all the thesis guidlines.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp]{revtex4-1}
\usepackage{graphicx} % Include figure files
\graphicspath{{figs/}}
\usepackage{color}
\usepackage{amsmath}
\usepackage{subfig}

\usepackage{nicefrac} % Include for inline fractions
\usepackage{url}

% \usepackage{cite}
\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage{color}  % Coloured text etc.
%
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}

\newcommand{\addcite}{{\bf \color{red} []}}
\newcommand{\xvec}{\mathbf{x}}
\newcommand{\thetavec}{\boldsymbol \theta}

\newcommand\V{\Phi}
\newcommand\Vk{\tilde\Phi(\vec k)}
\newcommand\volume{V}

\newcommand{\pink}[1]{{\bf \color{pink} #1}}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\orange}[1]{{\bf \color{orange} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

%\newcommandx{\davidsays}[1]{\begin{mdframed}[backgroundcolor=blue!24]\textbf{David says:} #1\end{mdframed} }
%\newcommandx{\davidsays}[2][1=inline]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{David says:} #2} }
\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }

%% \renewcommand\davidsays[1]{}
%% \renewcommand{\corysays}[2][1=inline]{}
%TC:macro \corysays [ignore]
%TC:macro \davidsays [ignore]
%TC:macro \jordan [ignore]

\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }

\newtheorem{limitation}{Limitation}

\begin{document}
\title{
An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}
% other ideas:
%Gas delivery via pressure-swing adsorption in nanoporous materials: setting storage targets}
% A thermodynamic upper bound to gas delivery via pressure-swing adsorption in nanoporous materials:}

\author{Jordan K. Pommerenck}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}
\author{Cory M. Simon}
\affiliation{School of Chemical, Biological, and Environmental Engineering,
  Corvallis, OR 97331}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  Due to their low volumetric energy density at ambient conditions, both hydrogen and natural gas are challenging to economically store onboard environmentally sustainable vehicles []. One strategy to densify these gases is to pack the fuel tank with a porous material [].
  Metal-organic frameworks are tunable, nanoporous materials with large internal surface areas and show considerable promise for densifying gases []. The US Department of Energy (DOE) has set volumetric deliverable capacity targets [] which, if met, would help to enable commercial adoption of hydrogen/natural gas as transportation fuels. Many have attempted to establish theoretical upper bounds on the deliverable capacity via pressure-swing adsorption using simplified models of the gas-substrate and gas-gas interaction []; none have established a rigorous upper bound on the deliverable capacity.
  Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing. % "pressure-swing storage doesn't make sense to me. storage and delivery  maybe but awk.
  % I added rigid since that could be important...
  To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
  \davidsays{Maybe we can add a phrase for an optimal uniform attraction?}
  Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
  We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
  % However, this upper bound directly leads to important realizations which should guide future development.
  However, limitations to the scope of applicability of our upper bound guide future material development efforts.
 % Firstly, one could extract the gas at a higher temperature than that used while filling [].
  Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes [].
  Secondly, the fundamental physics of our upper bound do not pertain to
  % rule out
  any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target [].
  % firstly, secondly vs first, second?
\end{abstract}

\newcommand\pfull{\ensuremath{p_{\text{full}}}}
\newcommand\pempty{\ensuremath{p_{\text{empty}}}}

\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}

\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}

\maketitle

\newcommand\gst{\ensuremath{\Delta g_{st}}}

\paragraph*{to do}
\begin{enumerate}
    \item See david/corysays
    \item make the code presentable
    \item do the FIXME in code
\end{enumerate}

\input{gas-ads-body.tex}

\bibliography{bibfile}

\input{gas-ads-appendix.tex}

\end{document}