### 2020-05-08 ising/ising.tex
fix some 1/t and edit abstract and stuff

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}
\usepackage{braket}

% define colors for comments
\definecolor{dark-gray}{gray}{0.10}
\definecolor{light-gray}{gray}{0.70}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
\begin{document}

\title{Flat-histogram method comparison on the 2D Ising model
}

\author{Jordan K. Pommerenck} \author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  We compare the convergence of several flat-histogram methods applied
  to the 2D Ising model, including the recently introduced stochastic
  approximation with a dynamic update factor (SAD) method.  We compare
  this method with the Wang-Landau (WL) method, the $1/t$ variant of
  the WL method, and standard stochastic approximation Monte Carlo
  (SAMC).  In addition, we consider a procedure WL followed by a
  ``production run'' with fixed weights that refines the estimation of
  the entropy.  To our knowledge, this work is the first to test this
  approach against other methods.
  %% To our knowledge, the
  %% convergence rate of this production-run WL approach has never been 
%% We examine the convergence of stochastic approximation with a dynamic
%% update factor (SAD) with other flat-histogram methods applied to the 2D
%% Ising model. The comparison with stochastic approximation Monte Carlo (SAMC)
%% and Wang-Landau (WL) methods shows that SAD performs robustly and without user
%% input knowledge of an energy range. We also compare SAD with WL followed by a
%% ``production run'' with fixed weights which is the first comparison of its kind
%% recorded.
We find that WL followed by a production run \emph{does} converge to
the true density of states, in contrast to pure WL.  Three of the
methods converge robustly: SAD, $1/t$-WL, and WL followed by a
production run.  Of these, SAD does not require \emph{a priori}
knowledge of the energy range.
\end{abstract}

\maketitle

\input{ising-body.tex}

\bibliography{ising} % Produces the bibliography via BibTeX.

\end{document}