### 2020-05-22 gas-adsorption/gas-ads.tex
made cory comment fixes for abstract.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp]{revtex4-1}
\usepackage{graphicx} % Include figure files
\graphicspath{{figs/}}
\usepackage{color}
\usepackage{amsmath}
\usepackage{subfig}

\usepackage{nicefrac} % Include for inline fractions
\usepackage{url}

% \usepackage{cite}
\usepackage{xargs}
% Use more than one optional parameter in a new commands
\usepackage{color}
%
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}

\newcommand{\addcite}{{\bf \color{red} []}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\kvec}{\mathbf{k}}
\newcommand{\thetavec}{\boldsymbol \theta}

\newcommand\V{\Phi}
\newcommand\Vk{\tilde\Phi(\kvec)}
\newcommand\volume{V}

\newcommand{\pink}[1]{{\bf \color{pink} #1}}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\orange}[1]{{\bf \color{orange} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }

\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }

\newtheorem{limitation}{Limitation}

\begin{document}
\title{
An upper bound to gas storage and delivery via pressure-swing adsorption in
porous materials}

\author{Jordan K. Pommerenck}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}
\author{Cory M. Simon}
\affiliation{School of Chemical, Biological, and Environmental Engineering,
  Corvallis, OR 97331}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
Due to their low volumetric energy density at ambient conditions, both hydrogen
and natural gas are challenging to economically store onboard vehicles as
fuels~\cite{mason2014evaluating, sircar2002pressure}. One strategy to densify
these gases is to pack the fuel tank with a porous
material~\cite{schoedel2016role}. Metal-organic frameworks (MOFs) are tunable,
nanoporous materials with large internal surface areas and show considerable
promise for densifying gases~\cite{makal2012methane,mason2014evaluating,
suh2011hydrogen,garcia2018benchmark, schoedel2016role}. The US Department of
Energy (DOE) has set volumetric deliverable capacity
targets~\cite{simon2015materials, h2targetsDOE} which, if met, would help
enable commercial adoption of hydrogen/natural gas as transportation fuels.
Many have attempted to establish theoretical upper bounds on the deliverable
capacity via pressure-swing adsorption using simplified models of the
gas-substrate and gas-gas interaction~\cite{gomez2014exploring,
gomez2017impact, kaija2018high, lee2019predicting}; none have established a
rigorous upper bound on the deliverable capacity. Here, we present a
theoretical upper bound on the deliverable capacity of a rigid material via an
isothermal pressure-swing. To provide an extremum, we consider a substrate that
provides a spatially uniform potential energy field for the gas. Our bound is
unique in that it directly relies on experimentally measured properties of the
(real) bulk gas, without making approximations. We conclude that the goals set
by the US DOE for natural gas and hydrogen storage are theoretically possible,
but sufficiently close to the upper bound as to be impractical for any real
porous material. However, limitations to the scope of applicability of our
upper bound guide future material development efforts. Firstly, one could heat
the adsorbent to drive off residual gas trapped in the
adsorbent~\cite{gomez2014exploring}. Secondly, the fundamental physics of our
upper bound do not pertain to any material that changes its structure in
response to adsorbed gas, suggesting that flexible materials could still
satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly,
mason2015methane}.

\end{abstract}

\newcommand\pfull{\ensuremath{p_{\text{full}}}}
\newcommand\pempty{\ensuremath{p_{\text{empty}}}}

\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}

\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}

\maketitle

\newcommand\gst{\ensuremath{\Delta g_{st}}}

\input{gas-ads-body.tex}

\bibliography{bibfile}

\appendix
\clearpage
\input{gas-ads-appendix.tex}

\end{document}