### 2020-05-27 dissertation/dissertation.tex
working on a new template for thesis.

\documentclass[double,12pt]{beavtex}
\usepackage{graphicx}
\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command

\renewcommand{\chaptername}{}

\title{Branching Random Walk and Probability Problems
from Physics and Biology}
\author{Torrey Johnson}
\degree{Doctor of Philosophy}
\doctype{Dissertation}
\department{Mathematics}
\depttype{Department}
\depthead{Chair}
\major{Mathematics}
\advisor{Edward C Waymire}
\submitdate{June 7, 2012}
\commencementyear{2013}
\abstract{\input{chapter-5.tex}}
\acknowledgements{\input{chapter-5.tex}}

\begin{document}

\maketitle

\mainmatter

\begingroup
\let\clearpage\relax
\begin{center}
\large
Branching Random Walk and Probability Problems
from Physics and Biology
\normalsize
\end{center}
\chapter{Introduction}
hello this is the chapter
%\input{introduction.tex}
\endgroup

% \chapter{Tree Polymers in the Infinite Volume Limit at Critical Strong Disorder}
% \input{csd.tex}

% \chapter{Simple Branching Random Walk and an Associated Tree Polymer}
% \input{brwtree.tex}

% \chapter{Branching for a Decision: A Site-Selection Model for Honeybee Swarms}
% \input{beetext.tex}

% \chapter{Conclusion}
% \input{conclusion.tex}

% \bibliographystyle{plain}
% \bibliography{bib}

% \appendix
% \chapter{Code for Swarm Simulation}
% \input{appendix.tex}
\end{document}