### 2020-05-27 dissertation/abstract.tex
added a new dissertation tex which conforms to OSU guidlines.

This work introduces the novel flat-histogram Monte Carlo (MC) method
stochastic approximation with a dynamic update factor (SAD) and explores the
convergence properties of a variety of related weight-based MC methods. The new
method is applied to a number of physical `test’ systems including the 2D Ising
model, a square-well fluid, and a Lennard-Jones cluster. A driving motivation
for developing novel Monte Carlo methods that have physically based tunable
parameters is to provide simulation methods for exploring metal-organic
frameworks (MOFs). A theoretical framework for gas adsorption and delivery in
metal-organic frameworks is developed and compared with experimental data. SAD
is introduced as a powerful method for examining thermodynamic properties for
MOFs. A preliminary groundwork is laid for the future development of a
multi-dimensional SAD which would be able to compute all properties of interest
for any given MOF thereby providing a powerful way for researchers to determine
any porous media's suitability for gas storage applications.