### 2020-05-27 dissertation/acknowledgements.tex

I gratefully acknowledge my graduate advisor David Roundy for his constant
support and guidance throughout my research. His breadth of knowledge and
hands-off approach to advising have tremendously helped my development as an
independent scientific researcher. I express thanks to my entire graduate
committee: David McIntyre, Davide Lazzati, Yun-Shik Lee, and graduate council
representative Chih-hung (Alex) Chang. I would like to acknowledge the many
outstanding undergraduate and graduate students that have positively impacted
me personally and professionally while at Oregon State University. Finally, I
would like to thank my parents for their unconditional support throughout my
Ph.D. research.