### 2020-05-27 dissertation/contributors.tex

Michael A. Perlin worked on much of the early C{}\verb!++! codebase. This
codebase was used primarily in the early study of many of the MC methods before
switching to RUST. In addition, he helped in the editing phase during the
preparation of the first manuscript \emph{Stochastic approximation Monte Carlo
with a dynamic update factor}. Tanner T. Simpson performed many of the early
simulations for the flat-histogram methods using the C{}\verb!++! codebase. He
also developed using Python some of the plotting scripts that allowed the saved
C{}\verb!++! data to be visualized. In addition, he helped in the editing phase
during the preparation of the first manuscript \emph{Stochastic approximation
Monte Carlo with a dynamic update factor}. Cory M. Simon collaborated together
with us on the third manuscript \emph{An upper bound to gas delivery via
pressure-swing adsorption in nanoporous materials}. He contributed
significantly to the introduction section of the paper and helped in the
editing phase throughout preparation. In addition, he designed the first figure
in the paper and the first figure in the supplemental to the paper.