### 2020-05-27 dissertation/dissertation.tex

\documentclass[double,12pt]{beavtex}

\usepackage{fancyhdr}
% link to section command
\newcommand{\HR}[1]{\textbf{\hyperref[#1]{\ref{#1}}}}

% for creating an index at end of document
\usepackage{imakeidx}
\makeindex[columns=1, title=Alphabetical Index]

% indexing commands
\newcommand{\bind}[1]{\textbf{#1}\index{#1}}
\newcommand{\iind}[1]{\textit{#1}\index{#1}}
\newcommand{\ind}[1]{#1\index{#1}}

\usepackage[hypertexnames=false]{hyperref}
\hypersetup{
    unicode=false,          % non-Latin characters in Acrobat’s bookmarks
    pdftoolbar=true,        % show Acrobat’s toolbar?
    pdfmenubar=true,        % show Acrobat’s menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={Pommerenck Thesis},    % title
    pdfauthor={Jordan K. Pommerenck},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={flat-histogram, Monte Carlo, Thesis}, % list of keywords
    pdfnewwindow=true,      % links in new PDF window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
    citecolor=black,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
%\usepackage{blindtext}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[document]{ragged2e}
% differential and partials package
\usepackage[thinc]{esdiff}
\usepackage{physics}
\newcommand{\D}[1]{\text{d}#1}
% the nth superscript package
\usepackage[super]{nth}

\usepackage{indentfirst}
\usepackage{graphicx}
\graphicspath{{../square-well/}{../ising/}{../gas-adsorption/figs/}}

\usepackage{caption}
%\usepackage{subcaption}
\usepackage{float}
\usepackage{amsmath,amssymb}
\usepackage{nicefrac}

% Use more than one optional parameter in a new commands
\usepackage{xargs}
\usepackage[pdftex,dvipsnames]{xcolor}
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}
\usepackage{braket}

% gas adsorption commands
\usepackage{subfig}
\usepackage{url}
%
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}

\newcommand{\addcite}{{\bf \color{red} []}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\kvec}{\mathbf{k}}
\newcommand{\thetavec}{\boldsymbol \theta}
\newcommand\V{\Phi}
\newcommand\Vk{\tilde\Phi(\kvec)}
\newcommand\volume{V}
\newcommand\pfull{\ensuremath{p_{\text{full}}}}
\newcommand\pempty{\ensuremath{p_{\text{empty}}}}
\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}
\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}
\newcommand\gst{\ensuremath{\Delta g_{st}}}

\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }

\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }

\newtheorem{limitation}{Limitation}

\renewcommand{\chaptername}{}

\title{Advancing Renewable Gas Storage using Flat-histogram Methods}
\author{Jordan K. Pommerenck}
\degree{Doctor of Philosophy}
\doctype{Dissertation}
\department{Physics}
\depttype{Department}
\depthead{Head}
\major{Physics}
\advisor{David Roundy}
\submitdate{June 8, 2020}
\commencementyear{2020}
\abstract{\input{abstract.tex}}
\acknowledgements{\input{acknowledgements.tex}}
\contributors{\input{contributors.tex}}

\begin{document}

\maketitle

\mainmatter

\begingroup
\let\clearpage\relax
\begin{center}
\large
Advancing Renewable Gas Storage using Flat-histogram Methods
\normalsize
\end{center}
\chapter{Introduction}
\input{introduction.tex}
\endgroup

\chapter{Stochastic Approximation Monte Carlo with a Dynamic Update
Factor}
\vspace*{2.0cm}

\textbf{Authors:} \\
Jordan K. Pommerenck, Tanner T. Simpson, Michael A. Perlin,
and David Roundy

\vspace*{4.0cm}

\textbf{Physical Review E} \\
\href{https://doi.org/10.1103/PhysRevE.101.013301}{Journal Online Access} \\
\textbf{101(1)}, 013301 (2020)~\cite{pommerenck2020stochastic}

\clearpage{}

\input{../square-well/sq-well-body.tex}

\chapter{Flat-histogram method comparison on 2D Ising Model
}
\input{../ising/ising-body.tex}

\chapter{An upper bound to gas storage and delivery via pressure-swing
adsorption in porous materials}
\input{../gas-adsorption/gas-ads-body.tex}

\chapter{Conclusion}
\input{conclusion.tex}

\bibliographystyle{plain}
\bibliography{thesis}

\appendix
\input{../gas-adsorption/gas-ads-appendix.tex}
\end{document}