### 2020-06-02 gas-adsorption/gas-ads.tex
final paper changes.

\RequirePackage[mathlines]{lineno}
\documentclass[letterpaper,twocolumn,amsmath,amssymb,aps,pre,10pt]{revtex4-1}
\setlength{\columnsep}{1cm}
\linenumbers

\usepackage{graphicx} % Include figure files
\graphicspath{{figs/}}
\usepackage{color}
\usepackage{amsmath}

%\usepackage{lineno} % I added this to do line numbering
\usepackage{subcaption}
\usepackage{nicefrac} % Include for inline fractions
\usepackage{url}
%\usepackage[hypertexnames=false]{hyperref}
% \hypersetup{
%     unicode=false,          % non-Latin characters in Acrobat’s bookmarks
%     pdftoolbar=true,        % show Acrobat’s toolbar?
%     pdfmenubar=true,        % show Acrobat’s menu?
%     pdffitwindow=false,     % window fit to page when opened
%     pdfstartview={FitH},    % fits the width of the page to the window
%     pdftitle={Pommerenck Thesis},    % title
%     pdfauthor={Jordan K. Pommerenck},     % author
%     pdfsubject={Subject},   % subject of the document
%     pdfcreator={Creator},   % creator of the document
%     pdfproducer={Producer}, % producer of the document
%     pdfkeywords={flat-histogram, Monte Carlo, Thesis}, % list of keywords
%     pdfnewwindow=true,      % links in new PDF window
%     colorlinks=true,       % false: boxed links; true: colored links
%     linkcolor=black,          % color of internal links (change box color with linkbordercolor)
%     citecolor=black,        % color of links to bibliography
%     filecolor=magenta,      % color of file links
%     urlcolor=cyan           % color of external links
% }

% \usepackage{cite}
\usepackage{xargs}
% Use more than one optional parameter in a new commands
\usepackage{color}
%
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}

\newcommand{\addcite}{{\bf \color{red} []}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\kvec}{\mathbf{k}}
\newcommand{\thetavec}{\boldsymbol \theta}

\newcommand\V{\Phi}
\newcommand\Vk{\tilde\Phi(\kvec)}
\newcommand\volume{V}

\newcommand{\pink}[1]{{\bf \color{pink} #1}}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\orange}[1]{{\bf \color{orange} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }

\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }

\newtheorem{limitation}{Limitation}

\begin{document}
\title{
An upper bound to gas storage and delivery via pressure-swing adsorption in
porous materials}

\author{Jordan K. Pommerenck}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}
\author{Cory M. Simon}
\affiliation{School of Chemical, Biological, and Environmental Engineering,
  Corvallis, OR 97331}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{linenomath}
\begin{abstract}
Due to their low volumetric energy density at ambient conditions, both hydrogen
and natural gas are challenging to economically store onboard vehicles as
fuels~\cite{mason2014evaluating, sircar2002pressure}. One strategy to densify
these gases is to pack the fuel tank with a porous
material~\cite{schoedel2016role}. Metal-organic frameworks (MOFs) are tunable,
nanoporous materials with large internal surface areas and show considerable
promise for densifying gases~\cite{makal2012methane,mason2014evaluating,
suh2011hydrogen,garcia2018benchmark, schoedel2016role}. The US Department of
Energy (DOE) has set volumetric deliverable capacity
targets~\cite{simon2015materials, h2targetsDOE} which, if met, would help
enable commercial adoption of hydrogen/natural gas as transportation fuels.
Many have attempted to establish theoretical upper bounds on the deliverable
capacity via pressure-swing adsorption using simplified models of the
gas-substrate and gas-gas interaction~\cite{gomez2014exploring,
gomez2017impact, kaija2018high, lee2019predicting}; none have established a
rigorous upper bound on the deliverable capacity. Here, we present a
theoretical upper bound on the deliverable capacity of a gas in a rigid
material via an isothermal pressure-swing. To provide an extremum, we consider
a substrate that provides a spatially uniform potential energy field for the
gas. Our bound is unique in that it directly relies on experimentally measured
properties of the (real) bulk gas, without making approximations. We conclude
that the goals set by the US DOE for room-temperature natural gas and hydrogen
storage are theoretically possible, but sufficiently close to the upper bound
as to be impractical for any real, rigid porous material. However, limitations
to the scope of applicability of our upper bound guide fuel tank design and
future material development. Firstly, one could heat the adsorbent to drive off
trapped, residual gas in the adsorbent~\cite{gomez2014exploring}. Secondly, the
physics of our upper bound do not pertain to any material that changes its
structure in response to adsorbed gas, suggesting that flexible materials could
still satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly,
mason2015methane}.

\end{abstract}
\end{linenomath}

\newcommand\pfull{\ensuremath{p_{\text{full}}}}
\newcommand\pempty{\ensuremath{p_{\text{empty}}}}

\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}

\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}

\maketitle

\newcommand\gst{\ensuremath{\Delta g_{st}}}

\input{gas-ads-body.tex}

\bibliography{bibfile}

\appendix
\clearpage
\input{gas-ads-appendix.tex}

\end{document}