### 2020-06-06 dissertation/abstract.tex
reformed my abstract in response to comments.

This work introduces the novel flat-histogram Monte Carlo (MC) method
stochastic approximation with a dynamic update factor (SAD) and explores the
convergence properties of a variety of related weight-based MC methods. The new
method is applied to a number of physical `test’ systems including the 2D Ising
model, a square-well fluid, and a 31 atom Lennard-Jones cluster. We find that
SAD performs robustly on all of the systems. Also, SAD efficiently samples the
entire energy space defined by a chosen temperature range (rather than using
unphysical parameters). Additionally, we develop a theoretical upper bound for
gas adsorption and delivery in porous materials and compare the upper bound with
experimental data.
A driving motivation for developing the novel Monte Carlo method SAD is to
provide a simulation method for computing the thermodynamic properties of
porous materials. By computationally determining a materials deliverable
capacity, researchers save countless development and experimental hours. In
particular, this body of research heavily contributes to enabling light-vehicle
gas storage applications.