### 2021-02-25 gas-adsorption/referee-response.md
start a response to referees

## Response to referees

# Referee 1

Referee 1 states that "Phenomenological predictions without any relation to real
materials are not very helpful, if not useless."  The referee fails to grasp
that our theory *does* take into account the gas, and places a fundamental limit
on the capacity of an entire class of materials (inflexible porous materials).
This is far from useless.  The referee suggests as superior an approach
involving screening hundreds of thousands of actual MOFs.  We will point out
that such an approach cannot possibly determine that there isn't a better MOF
out there, while our approach can do so.

TODO: We should find a place to edit the paper to make this more clear, and
state here that we did so.

TODO: Maybe include some more discussion of pore volume, just to illustrate that
our method does give information related to things like pore volume?

The other suggestion of Referee 1 is that we clearly state the capacity that we report, as discussed by Parilla et al.

TODO: We must do so.

# Referee 3

Referee 3 states that the conclusions are convincing, and makes two suggestions
for improving the paper:

1. The referee suggests that we attempt to quantify the possible effect of a
   flexible MOF on our upper bound.

TODO: Add a bit of discussion of how and why our model does not apply to
flexible MOFs, and why we cannot place *any* upper bound on a flexible MOF using
our approach.

2. The referee asks that we provide some guidelines in terms of working
   conditions to reach the DOE target.  They further suggest looking at a
   temperature difference to see how that could impact the deliverable capacity.

TODO: Add *some* sort of extra discussion on the topic.  Our model cannot
predict a temperature+pressure swing cycle because that would require
information about the entropy of adsorption at the two temperatures, which is
rather trickier, I think.